---
title: "Homework 3"
author: "Alyssa Monda"
date: "2/8/2017"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Rcmdr)
library(car)
library(RcmdrMisc)
library(stargazer)
library(caTools)
```


```{r}
# Import Data
View(White_wines)
```

Summary of data
```{r}
summary(White_wines)
```
Split Data into test and training sets. 
```{r}
# divide the dataset into a training and a testing set based on a random uniform number on fixed seed
set.seed(20170208) #sets the random seed you started at
White_wines$group <- runif(length(Galton$parent), min = 0, max = 1)
summary(White_wines) #generating uniform set based on the random number ranging between 0 and 1. length command asys make 905. this steps is creating a new variable and adding it to that dataset
White_wines.train <- subset(White_wines, group <= 0.90) #if number less than .9 they get assignmed to traning
White_wines.test <- subset(White_wines, group > 0.90) #if number is more than .9 they get assigned to the test group for the model
```



Another method?
```{r}
set.seed(12345)
White_wines$group= sample.split(12,SplitRatio=0.9)
head(White_wines)
train=subset(White_wines, White_wines$group==TRUE)
View(train)
test=subset(White_wines, White_wines$group==FALSE)
View(test)
```


Easier way? 
```{r}
train= White_wines [1:4408,] # this will put the first 90% of the rows into the training set
test= iris [4409:4898] # this will put the remaining 10% of the rows into the test set
```
####Neither work......

Scatterplot Matrix: quality, alcohol, chlorides, citric acid.


```{r}
scatterplotMatrix(~alcohol+chlorides+citric.acid+quality, reg.line=FALSE, 
  smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), 
  id.n=0, diagonal = 'density', data=White_wines)
```
Scatterplot Matrix: quality, density, fixed acidity, free sulfur dioxide. 

```{r}
scatterplotMatrix(~density+fixed.acidity+free.sulfur.dioxide+quality, 
  reg.line=FALSE, smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'density', data=White_wines)
```

Scatterplot Matrix: quality, pH, residual sugar, sulphates. 

```{r}
scatterplotMatrix(~pH+quality+residual.sugar+sulphates, reg.line=FALSE, 
  smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), 
  id.n=0, diagonal = 'density', data=White_wines)
```

Scatterplot Matrix: quality, total sulfur dioxide, volatile acidity.

```{r}
scatterplotMatrix(~quality+total.sulfur.dioxide+volatile.acidity, 
  reg.line=FALSE, smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'density', data=White_wines)
```
Correlation Matrices
```{r}
cor(White_wines[,c("alcohol","chlorides","citric.acid","density",
  "fixed.acidity","quality")], use="complete")
```
```{r}
cor(White_wines[,c("free.sulfur.dioxide","pH","quality","residual.sugar",
  "sulphates")], use="complete")
```
```{r}
cor(White_wines[,c("quality","total.sulfur.dioxide","volatile.acidity")], 
  use="complete")
```
#Variable           Correlation With quality
alcohol               0.435574715
chlorides            -0.209934411
citric.acid          -0.009209091
density              -0.307123313
fixed.acidity        -0.113662831
quality               1.000000000
free.sulfur.dioxide   0.008158067
pH                    0.099427246
residual.sugar       -0.097576829
sulphates             0.053677877
total.sulfur.dioxide -0.1747372
volatile.acidity     -0.1947230


Regression Model Alcohol and quality. 

```{r}
RegModel.Alcohol <- lm(alcohol~quality, data=White_wines)
summary(RegModel.Alcohol)
```
Coefficients:
            Estimate Std. Error t value Pr(>|t|)    
(Intercept)  6.95670    0.10626   65.47   <2e-16 ***
quality      0.60524    0.01788   33.86   <2e-16 ***

Residual standard error: 1.108 on 4896 degrees of freedom
Multiple R-squared:  0.1897,	Adjusted R-squared:  0.1896 
F-statistic:  1146 on 1 and 4896 DF,  p-value: < 2.2e-16

Regression Model fixed.acidity and quality.

```{r}
RegModel.fixed.acidity <- lm(fixed.acidity~quality, data=White_wines)
summary(RegModel.fixed.acidity)
```
Regression Model volatile.acidity and quality.
```{r}
RegModel.volatile.acidity <- lm(volatile.acidity~quality, data=White_wines)
summary(RegModel.volatile.acidity)
```
Regression Model citric.acid and quality.
```{r}
RegModel.citric.acid <- lm(citric.acid~quality, data=White_wines)
summary(RegModel.citric.acid)
```

Regression Model residual.sugar and quality.
```{r}
RegModel.residual.sugar <- lm(residual.sugar~quality, data=White_wines)
summary(RegModel.residual.sugar)
```

Regression Model chlorides and quality.
```{r}
RegModel.chlorides <- lm(chlorides~quality, data=White_wines)
summary(RegModel.chlorides)
```

Regression Model free.sulfur.dioxide and quality.
```{r}
RegModel.free.sulfur.dioxide <- lm(free.sulfur.dioxide~quality, data=White_wines)
summary(RegModel.free.sulfur.dioxide)
```

Regression Model total.sulfur.dioxide and quality.
```{r}
RegModel.total.sulfur.dioxide <- lm(total.sulfur.dioxide~quality, data=White_wines)
summary(RegModel.total.sulfur.dioxide)
```

Regression Model density and quality.
```{r}
RegModel.density <- lm(density~quality, data=White_wines)
summary(RegModel.density)
```

Regression Model pH and quality.
```{r}
RegModel.pH <- lm(pH~quality, data=White_wines)
summary(RegModel.pH)
```

Regression Model volatile.acidity and quality.
```{r}
RegModel.sulphates <- lm(sulphates~quality, data=White_wines)
summary(RegModel.sulphates)
```

Multiple Regression: Full Model 
```{r}
# let the regression rip
WhiteWine.multregmod1 <- lm(quality ~ alcohol + chlorides + sulphates + total.sulfur.dioxide + volatile.acidity + density + pH + fixed.acidity + free.sulfur.dioxide + citric.acid, data= White_wines)

#and see what we have
summary(WhiteWine.multregmod1)
```

```{r}

#added variable plots
avPlots(WhiteWine.multregmod1, id.n=2, id.cex=0.7)
#id.n - identify n most influential observations so you can pick out outlier values labeling them as farmers babysitters etc
#id.cex - controls the size of the dot
```


```{r}

# run the qq-plot
qqPlot(WhiteWine.multregmod1, id.n=3)
# here, id.n identifies the n observations with the largest residuals in absolute value
```

Residuals
```{r}
# diagnostics for the first model with 3 independent variables
residualPlots(WhiteWine.multregmod1)

```
Are there any outliers?

```{r}
 
#run Bonferroni test for outliers
outlierTest(WhiteWine.multregmod1)
```
```{r}
#make influence plot
influencePlot(WhiteWine.multregmod1, id.n=3)
```

```{r}
#test for heteroskedasticity
ncvTest(WhiteWine.multregmod1) #tests for non constant variance. 
```
```{r}
vif(WhiteWine.multregmod1)
#if higher than 4 we want to take variable out b/c it is not independent and highly correlates with something in there 
```



Multiple Regression: Reduced Model (Missing citric.acid and pH)
```{r}
# let the regression rip
WhiteWine.multregmod2 <- lm(quality ~ alcohol + chlorides + sulphates + total.sulfur.dioxide + volatile.acidity + density + fixed.acidity + free.sulfur.dioxide, data= White_wines)

#and see what we have
summary(WhiteWine.multregmod2)
```

Multiple Regression: Reduced Model (Alcohol, chlorides, sulphates, volatile.acidity, and density)
```{r}
# let the regression rip
WhiteWine.multregmod3 <- lm(quality ~ alcohol + chlorides + sulphates + volatile.acidity + density, data= White_wines)

#and see what we have
summary(WhiteWine.multregmod3)
```

Multiple Regression: Reduced Model (alcohol+ chlorides, sulphates volatile.acidity, and density)
```{r}
# let the regression rip
WhiteWine.multregmod3 <- lm(quality ~ alcohol+ sulphates + chlorides + volatile.acidity + density, data= White_wines)

#and see what we have
summary(WhiteWine.multregmod3)
```

```{r}
# diagnostics for the first model with 3 independent variables
residualPlots(WhiteWine.multregmod3)

```
Multiple Regression: Reduced Model (alcohol, volatile.acidity and sulphates)
```{r}
# let the regression rip
WhiteWine.multregmod4 <- lm(quality ~ alcohol + volatile.acidity + sulphates, data= White_wines)

#and see what we have
summary(WhiteWine.multregmod4)
```
Residuals
```{r}
# diagnostics for the first model with 3 independent variables
residualPlots(WhiteWine.multregmod4)

```


Which independent vaariables are independent from one another? 

```{r}
vif(WhiteWine.multregmod4)
#if higher than 4 we want to take variable out b/c it is not independent and highly correlates with something in there 
```


```{r results="asis"}
# compare the results of the two regression models
stargazer(WhiteWine.multregmod1, WhiteWine.multregmod4,title="Comparison of 2 Regression outputs",type="text",align=TRUE)
#can only be seen when knitting to html if you change type to text you can see the table now type=html or text or latek as options
```


No clue how to interpret..........
