---
title: "HW3_WhiteWines"
author: "Alyssa Monda"
date: "2/14/2017"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Rcmdr)
library(car)
library(RcmdrMisc)
library(stargazer)
library(caTools)
```

```{r}
library(readr)
White_wines <- read.table("~/Desktop/Big Data/Regression-1/White_wines.csv", header=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)
View(White_wines)
```

```{r}
# Import Data
#View(White_wines)
```

##Summary of Data 

Look at a summary of the data. 
```{r}
summary(White_wines)
```

This dataset is composed of 13 variables. The dependent variable of interest is quality. We will investigate the relationship between the remaining variables (fixed acidity, volatile acid, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, density, pH, sulphates, and alcohol) and quality.

##Univariate Analysis

Quality appears to be normally distributed with scores ranging from a minimum of 3 to a maximum of 9, with a mean score of of 5.88 and a median of 6.0. A boxplot of quality shows the potential of outliers. These should be considered when interpretting the remainder of the analysis.

Figure 1
```{r}
with(White_wines, Hist(quality, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```
Figure 2
```{r}
Boxplot( ~ quality, data=White_wines, id.method="y")
```

The distribiution of the remaining variables can be seen in the histograms below. Residual sugar, alcohol, and volatile acid appear to have right skewed distributions, while not perfect, the other variables appear to have a normal distribution. 

Figure 3
```{r}
with(White_wines, Hist(alcohol, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```

Figure 4
```{r}
with(White_wines, Hist(chlorides, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```
Figure 5
```{r}
with(White_wines, Hist(citric.acid, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```
Figure 6
```{r}
with(White_wines, Hist(density, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```
Figure 7
```{r}
with(White_wines, Hist(fixed.acidity, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```
Figure 8
```{r}
with(White_wines, Hist(free.sulfur.dioxide, scale="frequency", 
  breaks="Sturges", col="darkgray"))
```
Figure 9
```{r}
with(White_wines, Hist(pH, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```
Figure 10
```{r}
with(White_wines, Hist(residual.sugar, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```
Figure 11
```{r}
with(White_wines, Hist(sulphates, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```
Figure 12
```{r}
with(White_wines, Hist(total.sulfur.dioxide, scale="frequency", 
  breaks="Sturges", col="darkgray"))
```
Figure 13
```{r}
with(White_wines, Hist(volatile.acidity, scale="frequency", 
  breaks="Sturges", col="darkgray"))
```


##Multivariate Analysis
To begin investigating potential relationships scattlot matrices have been run below. 


Figure 13: Scatterplot Matrix: quality, alcohol, chlorides, citric acid.
```{r}
scatterplotMatrix(~alcohol+chlorides+citric.acid+quality, reg.line=FALSE, 
  smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), 
  id.n=0, diagonal = 'density', data=White_wines)
```
Figure 14: Scatterplot Matrix: quality, density, fixed acidity, free sulfur dioxide. 

```{r}
scatterplotMatrix(~density+fixed.acidity+free.sulfur.dioxide+quality, 
  reg.line=FALSE, smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'density', data=White_wines)
```

Figure 15: Scatterplot Matrix: quality, pH, residual sugar, sulphates. 

```{r}
scatterplotMatrix(~pH+quality+residual.sugar+sulphates, reg.line=FALSE, 
  smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), 
  id.n=0, diagonal = 'density', data=White_wines)
```

Figure 16: Scatterplot Matrix: quality, total sulfur dioxide, volatile acidity.

```{r}
scatterplotMatrix(~quality+total.sulfur.dioxide+volatile.acidity, 
  reg.line=FALSE, smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'density', data=White_wines)
```

Linear Correlation analysis shows: 

```{r}
cor(White_wines[,c("alcohol","chlorides","citric.acid","density",
  "fixed.acidity","quality")], use="complete")
```
```{r}
cor(White_wines[,c("free.sulfur.dioxide","pH","quality","residual.sugar",
  "sulphates")], use="complete")
```
```{r}
cor(White_wines[,c("quality","total.sulfur.dioxide","volatile.acidity")], 
  use="complete")
```

#Variable           Correlation With quality (r)
alcohol               0.435574715
chlorides            -0.209934411
citric.acid          -0.009209091
density              -0.307123313
fixed.acidity        -0.113662831
free.sulfur.dioxide   0.008158067
pH                    0.099427246
residual.sugar       -0.097576829
sulphates             0.053677877
total.sulfur.dioxide -0.1747372
volatile.acidity     -0.1947230

There seems to be a weak positive relationship between alcohol and quality. Density, chlorides, total sulfur dioxide, and volatile acid, seem to have the strongest negative correlations with quality.

#Linear Regressions
To further investigate potential relationships between quality and the variables linear regressions have been run below. 

Regression Model Alcohol and quality. 

```{r}
RegModel.Alcohol <- lm(alcohol~quality, data=White_wines)
summary(RegModel.Alcohol)
```

Regression Model fixed.acidity and quality.

```{r}
RegModel.fixed.acidity <- lm(fixed.acidity~quality, data=White_wines)
summary(RegModel.fixed.acidity)
```

Regression Model volatile.acidity and quality.
```{r}
RegModel.volatile.acidity <- lm(volatile.acidity~quality, data=White_wines)
summary(RegModel.volatile.acidity)
```

Regression Model citric.acid and quality.
```{r}
RegModel.citric.acid <- lm(citric.acid~quality, data=White_wines)
summary(RegModel.citric.acid)
```

Regression Model residual.sugar and quality.
```{r}
RegModel.residual.sugar <- lm(residual.sugar~quality, data=White_wines)
summary(RegModel.residual.sugar)
```

Regression Model chlorides and quality.
```{r}
RegModel.chlorides <- lm(chlorides~quality, data=White_wines)
summary(RegModel.chlorides)
```

Regression Model free.sulfur.dioxide and quality.
```{r}
RegModel.free.sulfur.dioxide <- lm(free.sulfur.dioxide~quality, data=White_wines)
summary(RegModel.free.sulfur.dioxide)
```

Regression Model total.sulfur.dioxide and quality.
```{r}
RegModel.total.sulfur.dioxide <- lm(total.sulfur.dioxide~quality, data=White_wines)
summary(RegModel.total.sulfur.dioxide)
```

Regression Model density and quality.
```{r}
RegModel.density <- lm(density~quality, data=White_wines)
summary(RegModel.density)
```

Regression Model pH and quality.
```{r}
RegModel.pH <- lm(pH~quality, data=White_wines)
summary(RegModel.pH)
```

Regression Model sulphates and quality.
```{r}
RegModel.sulphates <- lm(sulphates~quality, data=White_wines)
summary(RegModel.sulphates)
```

From this we see neither citric acid nor free sulfur dioxide appear to have a significant linear relationship with quality. 

#Multiple Regressions

Now we will begin building a model using multiple regressions. However, prior to building the model, we will first split our dataset into a training and testing set. 


```{r}
set.seed(20170214) #Random Number seed is the date
White_wines$group <- runif(length(White_wines$quality), min = 0, max = 1) #create a new variable to add to dataset to distribute random numbers from 0-1

White_wines.train <- subset(White_wines, group <= 0.90) #assign 90% of the data to the training set
White_wines.test <- subset(White_wines, group > 0.90) #assign remaining data to the test set

#Did it work?
summary(White_wines.train)
summary(White_wines.test)
```

Now we will begin with a full model including all variables. 
```{r}
LinearModel.Full <- lm(quality ~ alcohol + chlorides + citric.acid + 
  density + fixed.acidity + free.sulfur.dioxide + pH + residual.sugar 
  + sulphates + total.sulfur.dioxide + volatile.acidity, 
  data=White_wines.train)
summary(LinearModel.Full)
```
The full model can be used to explain 28% of the variability in taste. The F statistic is 156.9 and is highly significant. We will investigate what occurs as this model is reduced. 

To continue we will use the backwards selection strategy and remove all variable that were not significant in the full model. 

Reduced Model 1 will include alcohol, density, fixed acidity, free sulfur dioxide, pH, residal sugar, sulphates, volatile acidity. 

```{r}
LinearModel.2 <- lm(quality ~ alcohol +  density + fixed.acidity + 
  free.sulfur.dioxide +  pH + residual.sugar + sulphates +  volatile.acidity, 
  data=White_wines.train)
summary(LinearModel.2)
```
This reduced model can still be used to explain 28% of the variability in taste. The F statistic increased to 215.5 and is highly significant. We will investigate what occurs as this model is reduced.

View influential variables
Figure 17
```{r}
#added variable plots
avPlots(LinearModel.2, id.n=2, id.cex=0.7)
#id.n - identify n most influential observations so you can pick out outlier values labeling them as farmers babysitters etc
#id.cex - controls the size of the dot
```
Look at residuals

Figure 18
```{r}
# run the qq-plot
qqPlot(LinearModel.2, id.n=3)
# here, id.n identifies the n observations with the largest residuals in absolute value
```
Figure 19
```{r}
# diagnostics for the first model with 3 independent variables
residualPlots(LinearModel.2)
```
Outlier
```{r}
#run Bonferroni test for outliers
outlierTest(LinearModel.2)
```

Figure 20: Influence Plot
```{r}
#make influence plot
influencePlot(LinearModel.2, id.n=3)
```
Figure 21: Heteroskedascity
```{r}
#test for heteroskedasticity
ncvTest(LinearModel.2) #tests for non constant variance. 
```

```{r}
vif(LinearModel.2)
#if higher than 4 we want to take variable out b/c it is not independent and highly correlates with something in there 
```


Based on the previous plots/analysis we further reduce the model. will remove density, and residual sugar, and free sulfur dioxide from the analysis. 

```{r}
LinearModel.3 <- lm(quality ~ alcohol + fixed.acidity + 
  free.sulfur.dioxide +  pH + residual.sugar + sulphates +  volatile.acidity, 
  data=White_wines.train)
summary(LinearModel.3)
```
This reduced model can still be used to explain 27% of the variability in taste. The F statistic increased to 232.5 and is still highly significant. 
Figure 23
```{r}
# diagnostics for the first model with 3 independent variables
residualPlots(LinearModel.3)
```


We will investigate what occurs as this model is further  reduced by removing free sulfur dioxide. 

```{r}
LinearModel.4 <- lm(quality ~ alcohol + fixed.acidity +  pH + residual.sugar + sulphates + volatile.acidity, 
  data=White_wines.train)
summary(LinearModel.4)
```
This reduced model can still be used to explain 26% of the variability in taste. The F statistic increased to 265.3 and is still highly significant.

Figure 24
```{r}
# diagnostics for the first model with 3 independent variables
residualPlots(LinearModel.4)
```
```{r}
vif(LinearModel.4)
#if higher than 4 we want to take variable out b/c it is not independent and highly correlates with something in there 
```

Residual sugar

```{r}
LinearModel.5 <- lm(quality ~ alcohol + fixed.acidity + residual.sugar + sulphates + volatile.acidity, 
  data=White_wines.train)
summary(LinearModel.5)
```
This reduced model can still be used to explain 26% of the variability in taste. The F statistic increased to 317.2 and is still highly significant. An increase in the adjusted R squared indicates it may fit better than the previous model.

Figure 25
```{r}
# diagnostics for the first model with 3 independent variables
residualPlots(LinearModel.5)
```
```{r}
vif(LinearModel.5)
#if higher than 4 we want to take variable out b/c it is not independent and highly correlates with something in there 
```
Based on the residuals I would like to see what happens when residual sugar and fixed acidity are removed from the model.

```{r}
LinearModel.5 <- lm(quality ~ alcohol + sulphates + volatile.acidity, 
  data=White_wines.train)
summary(LinearModel.5)
```
This model does not appear to fit better than the previous model. It only accounts for 24% of the variability and while still significant the adjust R squared value has decreased from .26 to .23. 

```{r}
compareCoefs(LinearModel.2, LinearModel.3, LinearModel.4, LinearModel.5, 
  LinearModel.Full)
```
```{r results="asis"}
# compare the results of the two regression models
stargazer(LinearModel.3,LinearModel.4, LinearModel.5,title="Comparison of Regression outputs",type="text",align=TRUE)
#can only be seen when knitting to html if you change type to text you can see the table now type=html or text or latek as options
```

```{r}
#test for heteroskedasticity
ncvTest(LinearModel.4) #tests for non constant variance. All biomarkers fail this test. since p is big its a homoskedastic set
```
```{r}
vif(LinearModel.4)
#if higher than 4 we want to take variable out b/c it is not independent and highly correlates with something in there 
```
Figure 26
```{r}
#make influence plot
influencePlot(LinearModel.4, id.n=3)
```

Based on this data I believe LinearModel.4 to be the best model of this data. Currently the model accounts for 26% of the variability in the score for quality. While I am not pleased with the plots of the residuals or influential points, and I would also like to include less variables in the model. However I am unsure how much it is aceptable to balance these flaws for the amount of variability accounted for by the model. 

##Testing the Model
We will now run the model on the testing dataset to see if it is a good predictor of the values. 

```{r}
predict(LinearModel.4, White_wines.test)
```

  
```{r}
LinearModel.test <- lm(quality ~ alcohol + fixed.acidity +  pH + residual.sugar + sulphates + volatile.acidity, 
  data=White_wines.test)
summary(LinearModel.test)
```


Running this model on the test data provides a significant F statistic, with the model explaining 30% of the variability in the score for quality. This model uses alcohol, fixed acidity, pH, residual sugar, sulphates, volatile acidity to explain quality. This equation for this model is:

Y = 4.10 + (0.34)x1 + (-0.17)x2 + (-0.07)x3 + (0.01)x4 + (0.25)x5 + (-2.20)x6 + E

Where:
Y= quality
x1= alcohol
x2= fixed acidity
x3= pH
x4= residual sugar
x5= sulphates
x6= volatile acid
E= Error

Using this model it appears volatile acid influences quality the most. When keeping the other variables constant a 1 point change in volatile acid will cause a -2.20 change in the quality score of the wine. 

##Conclusion
Running the model on the full data set uses alcohol, fixed acidity, pH, residual sugar, sulphates, volatile acidity to explain quality to explain 26% of variability in quality score of the wine. Investigating the usefulness of model can be completed using the diagnositc test seen in the plots below. 

```{r}
LinearModel.testfull <- lm(quality ~ alcohol + fixed.acidity +  pH + residual.sugar + sulphates + volatile.acidity, 
  data=White_wines)
summary(LinearModel.testfull)
```
Figure 27
```{r}
# diagnostics for the first model with 3 independent variables
residualPlots(LinearModel.testfull)

```

Figure 28
```{r}
# diagnostics for the second model with 2 independent variables
residualPlots(LinearModel.testfull)
```

Figure 29
```{r}

#added variable plots
avPlots(LinearModel.testfull, id.n=2, id.cex=0.7)
#id.n - identify n most influential observations so you can pick out outlier values labeling them as farmers babysitters etc
#id.cex - controls the size of the dot
```

Figure 30
```{r}

# run the qq-plot
qqPlot(LinearModel.testfull, id.n=3)
# here, id.n identifies the n observations with the largest residuals in absolute value
```


```{r}
 
#run Bonferroni test for outliers
outlierTest(LinearModel.testfull)
```

Figure 31
```{r}

#identify highly influential points
influenceIndexPlot(LinearModel.testfull, id.n=3)
```

Figure 32
```{r}
#make influence plot
influencePlot(LinearModel.testfull, id.n=3)
```


```{r}
#test for heteroskedasticity
ncvTest(LinearModel.testfull) #tests for non constant variance. All biomarkers fail this test. since p is big its a homoskedastic set
```



```{r}
vif(LinearModel.testfull)
#if higher than 4 we want to take variable out b/c it is not independent and highly correlates with something in there 
```


The rmarkdown used to create this file can be found at <https://github.com/amonda/Regression-1>. The file is named New.Rmd.