<!-- R Commander Markdown Template -->

Replace with Main Title
=======================

### Your Name

### `r as.character(Sys.Date())`

```{r echo=FALSE}
# include this code chunk as-is to set options
knitr::opts_chunk$set(comment=NA, prompt=TRUE, out.width=750, fig.height=8, fig.width=8)
library(Rcmdr)
library(car)
library(RcmdrMisc)
```


```{r echo=FALSE}
# include this code chunk as-is to enable 3D graphs
library(rgl)
knitr::knit_hooks$set(webgl = hook_webgl)
```

Load data into R. 

```{r}
load("/Users/alyssamonda/Desktop/Big Data/Regression-1/White_wines.csv")
```
Make dataset active.

```{r}
names(White_wines) <- make.names(names(White_wines))
```

View summary statistics. 

```{r}
summary(White_wines)
```
Scatterplot Matrix: quality, alcohol, chlorides, citric acid.


```{r}
scatterplotMatrix(~alcohol+chlorides+citric.acid+quality, reg.line=FALSE, 
  smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), 
  id.n=0, diagonal = 'density', data=White_wines)
```
Scatterplot Matrix: quality, density, fixed acidity, free sulfur dioxide. 

```{r}
scatterplotMatrix(~density+fixed.acidity+free.sulfur.dioxide+quality, 
  reg.line=FALSE, smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'density', data=White_wines)
```

Scatterplot Matrix: quality, pH, residual sugar, sulphates. 

```{r}
scatterplotMatrix(~pH+quality+residual.sugar+sulphates, reg.line=FALSE, 
  smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), 
  id.n=0, diagonal = 'density', data=White_wines)
```

Scatterplot Matrix: quality, total sulfur dioxide, volatile acidity.

```{r}
scatterplotMatrix(~quality+total.sulfur.dioxide+volatile.acidity, 
  reg.line=FALSE, smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'density', data=White_wines)
```

Regression Model Alcohol and quality. 

```{r}
RegModel.Alcohol <- lm(alcohol~quality, data=White_wines)
summary(RegModel.Alcohol)
```
Coefficients:
            Estimate Std. Error t value Pr(>|t|)    
(Intercept)  6.95670    0.10626   65.47   <2e-16 ***
quality      0.60524    0.01788   33.86   <2e-16 ***

Residual standard error: 1.108 on 4896 degrees of freedom
Multiple R-squared:  0.1897,	Adjusted R-squared:  0.1896 
F-statistic:  1146 on 1 and 4896 DF,  p-value: < 2.2e-16

Regression Model fixed.acidity and quality.

```{r}
RegModel.fixed.acidity <- lm(fixed.acidity~quality, data=White_wines)
summary(RegModel.fixed.acidity)
```
Regression Model volatile.acidity and quality.
```{r}
RegModel.volatile.acidity <- lm(volatile.acidity~quality, data=White_wines)
summary(RegModel.volatile.acidity)
```
Regression Model citric.acid and quality.
```{r}
RegModel.citric.acid <- lm(citric.acid~quality, data=White_wines)
summary(RegModel.citric.acid)
```

Regression Model residual.sugar and quality.
```{r}
RegModel.residual.sugar <- lm(residual.sugar~quality, data=White_wines)
summary(RegModel.residual.sugar)
```

Regression Model chlorides and quality.
```{r}
RegModel.chlorides <- lm(chlorides~quality, data=White_wines)
summary(RegModel.chlorides)
```

Regression Model free.sulfur.dioxide and quality.
```{r}
RegModel.free.sulfur.dioxide <- lm(free.sulfur.dioxide~quality, data=White_wines)
summary(RegModel.free.sulfur.dioxide)
```

Regression Model total.sulfur.dioxide and quality.
```{r}
RegModel.total.sulfur.dioxide <- lm(total.sulfur.dioxide~quality, data=White_wines)
summary(RegModel.total.sulfur.dioxide)
```

Regression Model density and quality.
```{r}
RegModel.vdensity <- lm(density~quality, data=White_wines)
summary(RegModel.density)
```

Regression Model pH and quality.
```{r}
RegModel.pH <- lm(pH~quality, data=White_wines)
summary(RegModel.pH)
```

Regression Model volatile.acidity and quality.
```{r}
RegModel.sulphates <- lm(sulphates~quality, data=White_wines)
summary(RegModel.sulphates)
```


```{r}
cor(White_wines[,c("alcohol","chlorides","citric.acid","density",
  "fixed.acidity","quality")], use="complete")
```

```{r}
cor(White_wines[,c("free.sulfur.dioxide","pH","quality","residual.sugar",
  "sulphates")], use="complete")
```



```{r}
cor(White_wines[,c("quality","total.sulfur.dioxide","volatile.acidity")], 
  use="complete")
```

